### 2023-03-05 main/chapter_2ljets.tex
amenable
variables postit

Any event used in regions of the $\twoljets$-electroweak analysis is required
to have exactly two baseline leptons and exactly two signal leptons, meaning
that the event has two reconstructed leptons that pass both sets of criteria.
Requiring exactly two baseline leptons ensures that we veto events with soft
or low quality third leptons, thus ensuring that our data do not overlap with
other \atlas\ analysis that use three-lepton events.
Enforcing no overlap in this way makes the analysis more independent, and
therefore more cleanly combined in future meta-analyses.

@@

\subsection{Cheat-sheet}
\label{sec:2ljets_cheatsheet}
Uncomfortably many event variables are used to define the regions of this
analysis.
For future reference, their meanings, definitions, and references for these
variables are all collected here.
All dimensionful quantities are always reported in $\eV[G]$ units in this
thesis.
\begin{itemize}
\item $\met$: missing transverse momentum
\item $\metsig$: object-based missing transverse momentum significance,\\
described in Section~\ref{sec:2ljets_metsig}
\item $\ptjone$: transverse momentum of the hardest jet%
\vspace{0.5em}
\item $\mll$: invariant mass of the hardest two leptons
\item $\mjj$: invariant mass of the two hardest jets
\item $\mbb$: invariant mass of the two hardest $b$-tagged jets
\item $\mjetone$: mass of the hardest jet
\item $\mttwoll$: `stransverse' mass of the lepton pair with massless
invisibles,\\
described in Section~\ref{sec:2ljets_mt2}
\vspace{0.5em}
\item $\rll$: distance in $\eta\textrm{--}\phi$ between the lepton pair
\item $\rjj$: distance in $\eta\textrm{--}\phi$ between the two hardest jets%
\vspace{0.5em}
\item $\njet$: number of signal jets
\item $\nbtag$: number of b-tagged jets,
described in Section~\ref{sec:2ljets_btagging}
\vspace{0.5em}
\item $\dphillmet$: azimuthal angle between $\vec p_{\ell\ell}$ and $\ptmiss$,
where $\vec p_{\ell\ell}$ is the vector sum of the momenta of the leptons in
the lepton pair.
\item $\dphijmet$: azimuthal angle between the hardest jet and $\ptmiss$
\end{itemize}
All of these are physical properties of particles or collections of particles
that are reconstructed from data.
Note, however, that these are not identical to the same quantities that could
be defined from hypothetically true states of those particles ---
the reconstruction for a given `truth' state can vary with noise.
While I admit to dropping this distinction in lax language, it is needed to
explain the broad shapes of noisy variables such as $\mjj$ and indeed the
existence of $\metsig$, which relies on estimates of the reconstruction noise
and is ill-defined at a particle-truth level.

@@

Diboson backgrounds and C1N2 signals have more in common than just $\met$;
both also produce resonant $Z\to \ell\ell$ pairs
and do not enhance the rate of heavy-flavour jets, as displayed in
Figure~\ref{fig:2ljets_high_mll_b}.
Since signals and backgrounds have such similar shapes in $\mll$ and $\nbtag$,
sensitivity is only decreased by rejecting portions of these distributions.
Although most $\twoljets$-electroweak regions use tighter selection on these
variables to reduce backgrounds, this observation allows us to win
additional sensitivity in High regions by loose requirements
of $\mll \in 71\textrm{--}111\,\eV[G]$%
\footnote{%
We use interval notation in which
en-dashes `$a\textrm{--}b$' indicate open intervals $(a, b)$, and
concatenated intervals `$a\textrm{--}b\textrm{--}c$' indicate binning
with boundaries at $a$, $b$, and $c$.
This notation is consistent with the $\twoljets$
paper~\cite{atlas2022searches}.
}
and $\nbtag \leq 1$.
This $\mll$ window loosely selects the $m(Z)$ peak to scoop up nearby rare or
mismeasured events away from the central peak.
Similarly, allowing up to one $b$-tag improves yields without boosting
backgrounds, since the $\metsig$ requirement already removes most $\ttbar$
processes.

>>>>

Any event used in regions of the $\twoljets$-electroweak analysis is required
to have exactly two baseline leptons and exactly two signal leptons, meaning
that the event has two reconstructed leptons that pass both sets of criteria.
Requiring exactly two baseline leptons ensures that we veto events with soft
or low quality third leptons, thus ensuring that our data do not overlap with
other \atlas\ analysis that use three-lepton events.
Enforcing no overlap in this way makes the analysis more independent, and
therefore amenable to cleaner combinations in future meta-analyses.

@@

\subsection{Cheat-sheet}
\label{sec:2ljets_cheatsheet}
Uncomfortably many event variables are used to define the regions of this
analysis.
For future reference, their meanings, definitions, and references for these
variables are all collected here.
All dimensionful quantities are always reported in $\eV[G]$ units in this
thesis.
This list is roughly sorted to have the most important variables at the top.
\begin{itemize}
\item $\met$: missing transverse momentum
\item $\metsig$: object-based missing transverse momentum significance,\\
described in Section~\ref{sec:2ljets_metsig}
\item $\ptjone$: transverse momentum of the hardest jet%
\vspace{0.5em}
\item $\mll$: invariant mass of the hardest two leptons
\item $\mjj$: invariant mass of the two hardest jets
\item $\mbb$: invariant mass of the two hardest $b$-tagged jets
\item $\mjetone$: mass of the hardest jet
\item $\mttwoll$: `stransverse' mass of the lepton pair with massless
invisibles,\\
described in Section~\ref{sec:2ljets_mt2}
\vspace{0.5em}
\item $\rll$: distance in $\eta\textrm{--}\phi$ between the lepton pair
\item $\rjj$: distance in $\eta\textrm{--}\phi$ between the two hardest jets%
\vspace{0.5em}
\item $\njet$: number of signal jets
\item $\nbtag$: number of b-tagged jets,
described in Section~\ref{sec:2ljets_btagging}
\vspace{0.5em}
\item $\dphillmet$: azimuthal angle between $\vec p_{\ell\ell}$ and $\ptmiss$,
where $\vec p_{\ell\ell}$ is the vector sum of the momenta of the leptons in
the lepton pair.
\item $\dphijmet$: azimuthal angle between the hardest jet and $\ptmiss$
\end{itemize}
All of these are physical properties of particles or collections of particles
that are reconstructed from data.
Note, however, that these are not identical to the same quantities that could
be defined from hypothetically true states of those particles ---
the reconstruction for a given `truth' state can vary with noise.
While I admit to dropping this distinction in lax language, it is needed to
explain the broad shapes of noisy variables such as $\mjj$ and indeed the
existence of $\metsig$, which relies on estimates of the reconstruction noise
and is ill-defined at a particle-truth level.

Broadly speaking, some of these variables select necessary conditions for
signal-like events, and others help to improve signal purity within those
selection requirements.
Specifically, our signals produce $\met$, lepton pairs with
$\mll \approx m(Z)$, and jets leading to positive $\njet$.
Requiring appropriate conditions on these variables rejects large
Standard Model background yields without harming signal efficiency.
Other variables help to reduce background contamination with some trade-off
against signal efficiency; among these, $\metsig$ is the most generally strong
separating variable, and other variables are chosen in contexts where
they aid the local designs that we detail in Section~\ref{sec:2ljets_design}.

@@

Diboson backgrounds and C1N2 signals have more in common than just $\met$;
both also produce resonant $Z\to \ell\ell$ pairs
and do not enhance the rate of heavy-flavour jets, as displayed in
Figure~\ref{fig:2ljets_high_mll_b}.
Since signals and backgrounds have such similar shapes in $\mll$ and $\nbtag$,
sensitivity is only decreased by rejecting portions of these distributions.
Although most $\twoljets$-electroweak regions use tighter selection on these
variables to reduce backgrounds, this observation allows us to win
additional sensitivity in High regions by loose requirements
of $\mll \in 71\textrm{--}111\,\eV[G]$%
\footnote{%
We use interval notation in which
en-dashes `$a\textrm{--}b$' indicate open intervals $(a, b)$, and
concatenated intervals `$a\textrm{--}b\textrm{--}c$' indicate binning
with boundaries at $a$, $b$, and $c$.
This notation is consistent with the $\twoljets$
paper~\cite{atlas2022searches}.%
}
and $\nbtag \leq 1$.
This $\mll$ window loosely selects the $m(Z)$ peak to scoop up nearby rare or
mismeasured events away from the central peak.
Similarly, allowing up to one $b$-tag improves yields without boosting
backgrounds, since the $\metsig$ requirement already removes most $\ttbar$
processes.
